---
title: "Bayesian Hierarchical Model"
author: "Joseph Corritone"
format: html
editor: visual
---

## Bayesian Hierarchical Model for March 2026 Madness

This write-up covers the ideas, process, and analysis behind the development of a prediction model of the greatest sports tournament of the year, for submission into the University of Wisconsin-Madison Sports Analytics Club 2026 March Madness Data Challenge.

**Overview**

This will be a three-model development, starting with a base model, then incorporating efficiency & team characteristic metrics, and finally adding a dynamic element to the model.

```{r}
#| label: Preliminaries
#| warning: false
#| message: false

# install.packages(c("tidyverse", "brms", "tidybayes"))

library(tidyverse)
library(brms)
library(tidybayes)
```

### **THE BASE MODEL**

**2026 Regular Season & Conference Tournament Game Data**

```{r}
#| label: Regular Season Game Data

# Import Men's College Basketball regular season game data to build the base dataset:

season <- 2026    # Change depending on which seasons to include data from

regszn_compact_results <- read_csv("MRegularSeasonCompactResults.csv") %>%
  filter(Season == season) %>%
  select(Season, WTeamID, WScore, LTeamID, LScore, WLoc)

game_winners <- regszn_compact_results %>%
  transmute(Season, team = WTeamID, opponent = LTeamID, team_win = 1,
            team_score = WScore, opponent_score = LScore,
            home = if_else(WLoc == "H", 1, 0),
            neutral = if_else(WLoc == "N", 1, 0))

game_losers <- regszn_compact_results %>%
  transmute(Season, team = LTeamID, opponent = WTeamID, team_win = 0,
            team_score = LScore, opponent_score = WScore,
            home = case_when(WLoc == "A" ~ 1, TRUE ~ 0),
            neutral = if_else(WLoc == "N", 1, 0))

compact_game_results <- bind_rows(game_winners, game_losers) %>%
  mutate(score_diff = team_score - opponent_score)

```

```{r}
#| label: Team Ratings

# Methodology for developing team ratings in the base model. Will be developed further in future modeling



```
